As-you-go" instead of "after-the-fact": A network approach to scholarly communication and evaluation

### 2018-03-30 docs/manuscript.Rmd
Revise network structure

  Research outputs are typically papers, which report on at least one research cycle after it has occurred. The communicative design of papers embeds hindsight and its biases in the reporting of results. Moreover, this design eliminates the verification of the chronology within a paper. On the other hand, the paper encompasses so much that citations to other papers can indicate a tangent or a crucial link. Additionally, the paper is a bottleneck for what is communicated: It cannot properly deal with code, data, materials, etc.

  When stages of research are communicated separately and as they occur, it changes the communicative design to eliminate hindsight and allows more types of outputs to be communicated as separate modules. For example, a theory can be communicated first and hypotheses communicated afterwards, as a direct descendant of the theory. Subsequently, a design can be linked as a direct descendant of the hypotheses, materials as a direct descendant of the design, and so on. This would allow for the incorporation of materials, data, and analysis code (amongst others). In this structure, many nodes could link to a single node (e.g., replication causes many data nodes to connect to the same hypotheses node) but one node can also link to many other nodes (e.g., when hypotheses follow from multiple theories or when a meta-analytic node is linked to many results nodes).

  Figure 1 shows a simple example of how these different modular research outputs (i.e., nodes) would directly connect to each other. The connection between these nodes only shows the direct descendance and could still include citations to other pieces of information. For example, a discussion section could be a direct descendant of a results section and could still include citations to other relevant findings. When one research cycle ends, a new one can link to the last node, continuing the chain of descendance. Incorporating the direct descendancy of these knowledge modules builds a different kind of network than citation and authorship networks. As such, this network would be an addition to these already existing citation and authorship networks; it does not seek to replace them.

![**Figure 1. A Directed Acyclic Graph (DAG) of connected research stages. The ordering is chronological (top-bottom) and therefore nodes that are situated below one another cannot refer upwards.**](C:/Users/s379011/Dropbox/projects/2017network-rankings/docs/Topological_Ordering.png)

  Given that these modular outputs would be communicated as they occur, chronology is directly embedded in the communication process with many added benefits. For example, preregistration of hypotheses tries to ensure that predictions precede observations, which would be embedded with piecemeal communication where predictions are communicated when they are made [exploratory research could be communicated without hypotheses; for a more extensive discussion of the benefits and limits of preregistration see @10.1073/pnas.1708274114]. Moreover, if modular research outputs are communicated as they are produced, selective reporting (i.e., publication bias) is reduced by having already communicated the data before results are generated.

  With immutable append-only registers, the chronology and content integrity of these outputs can be ensured and preserved over time. This can occur efficiently and elegantly with the Dat protocol. In short, the Dat protocol is a peer-to-peer protocol (i.e., decentralized and openly accessible) that provides non-adjustable timestamps to each change that occurs within a folder, which is given a permanent unique address on the peer-to-peer Web (36^64 addresses possible) [@Dat2017]. The full details, implications, and potential implementations of this protocol for scholarly communication fall outside of the scope of this paper [see @dat-com].

@@

  With a chronological ordering of various research outputs and their parent relations, a directional adjacency matrix can be extracted for network analysis. Table 1 shows the directional adjacency matrix for Figure 1. Parent nodes must precede the child nodes in time, therefore only $\frac{J(J-1)}{2}$ of cells of the adjacency matrix are filled in, where $J$ is the number of research outputs.

@@

  We propose to communicate research in piecemeal "as-you-go" units (e.g., theory followed by hypotheses, etc.) instead of large "after-the-fact" papers. Piecemeal communication opens up the possibility of a network of knowledge to come into existence when these pieces are linked (e.g., results descend from data). This network of knowledge would be supplementary to traditional citation networks and could facilitate new evaluation tools that are based in the question of interest than generic "one size fits all" evaluation indicators (e.g., Journal Impact Factor, citation counts, number of publications). Given the countless questions and operationalizations possible to evaluate research in a network of knowledge, we hope this would increase the focus on indicators as a tool in the evaluation process instead of indicators being the evaluation process itself [@Hicks2015].

@@

  Communicating scholarly research in smaller "as-you-go" units might also address other threats to research sustainability. In piecemeal "as-you-go" communication, selective publication based on results would be reduced because data would be communicated before results are known. Similarly, adjusting predictions after results are known would be reduced because predictions would be communicated before data are available (i.e., preregistration by design). Replications (or reanalyses) would be encouraged both for the replicated (the replicated node gets more child nodes, increasing its centrality) and the replicator (time investment is lower due to only having to add a data node that is linked to the materials node of the replicated). Self-plagiarism could be reduced by not forcing researchers to rehash the same theory across papers that spawn various predictions and studies. These various issues (amongst other out of scope issues) could be addressed jointly instead of each issue vying for importance for researchers, funders, or policy makers (amongst others).

  To encourage culture- and behavioral change, "after-the-fact" papers and piecemeal "as-you-go" outputs could co-exist (initially) and would not require researchers to make a zero-sum decision. Copyright is often transferred to publishers upon publication (resulting in pay-to-access), but only after a legal contract is signed. Hence, preprints cannot be legally restricted by publishers when they precede a copyright transfer agreement. However, preprints face institutional and social opposition [@Kaiser2017], where preprinting could exclude a manuscript for publication depending on editorial policies or due to fears of non-publication or scooping (itself a result of hypercompetition). In recent years, preprints have become more widely accepted and less likely to exclude manuscript publication (e.g., Science accepts preprinted manuscripts) [@Berg2017]. Similarly, sharing piecemeal "as-you-go" outputs would not be legally restricted by publishers and could ride the wave of preprint acceptance, but might also face institutional or social counterchange similar to the history of preprints. Researchers could communicate "as-they-go" and compile "after-the-fact" papers, facilitating co-existence and minimizing negative effects on career opportunities. Additionally, “as-you-go” modules could be used in scholarly fields where the provenance of information is important to findings.

  As far as we know, piecemeal "as-you-go" scholarly communication infrastructure has not yet been available to researchers in a sustainable way. The only thought style that has facilitated "as-you-go" reporting in the past decade is that of Open Notebook Science (ONS) [@Bradley2007], where researchers share their day-to-day notes and thoughts. However, ONS has remained on the fringes of the Open Science thought style and has not matured, limiting its usefulness and uptake. For example, ONS increases user control because communication occurs on personal domains, but does not have a mechanism of preserving the content. Considering reference rot occurs in seven out of ten scholarly papers containing Weblinks [@Klein2014], concern for sustainable ONS is warranted without further development of content integrity. Moreover, ONS increases information output without providing more possibilities of discovering that content.

>>>>

  Research outputs are typically research papers, which report on at least one research cycle after it has occurred. The communicative design of papers embeds hindsight and its biases in the reporting of results by being inherently reconstructive. Moreover, this design eliminates the verification of the chronology within a paper. On the other hand, the paper encompasses so much that citations to other papers can indicate a tangent or a crucial link. Additionally, the paper is a bottleneck for what is communicated: It cannot properly deal with code, data, materials, etc.

  When stages of research are communicated separately and as they occur, it changes the communicative design to eliminate hindsight and allows more types of outputs to be communicated as separate modules. For example, a theory can be communicated first and hypotheses communicated second, as a direct descendant of the theory. Subsequently, a study design can be linked as a direct descendant of the hypotheses, materials as a direct descendant of the design, and so on. This would allow for the incorporation of materials, data, and analysis code (amongst others). In this structure, many modules could link to a single module (e.g., replication causes many data modules to connect to the same hypotheses module) but one module can also link to many other modules (e.g., when hypotheses follow from multiple theories or when a meta-analytic node is linked to many results nodes).

  Figure 1 shows a simple example of how these different modular research outputs (i.e., nodes) would directly connect to each other. The connection between these nodes only shows the direct descendance and could still include citations to other pieces of information. For example, a discussion node could be a direct descendant of a results node and could still include citations to other relevant findings. When one research cycle ends, a new one can link to the last node, continuing the chain of descendance. Incorporating the direct descendancy of these knowledge modules builds a different kind of network than citation and authorship networks. As such, this network would be an addition to these already existing citation and authorship networks; it does not seek to replace them.

![**Figure 1. Two Directed Acyclic Graphs (DAGs) of connected research stages. The ordering is chronological (top-bottom) and therefore nodes that are situated below one another cannot refer upwards. Panel A shows a less complex network of modules; Panel B shows a more extensive network of modules.**](Topological_Ordering.png)

  Given that these modular outputs would be communicated as they occur, chronology is directly embedded in the communication process with many added benefits. For example, preregistration of hypotheses tries to ensure that predictions precede observations, which would be embedded with modular communication where predictions are communicated when they are made [exploratory research could be communicated without hypotheses; for a more extensive discussion of the benefits and limits of preregistration see @10.1073/pnas.1708274114]. Moreover, if modular outputs are communicated as they are produced, selective reporting (i.e., publication bias) is reduced by having already communicated the data before results are generated.

  With immutable append-only registers, the chronology and content integrity of these outputs can be ensured and preserved over time. This can occur efficiently and elegantly with the Dat protocol (without a blockchain; [perma.cc/GC8X-VQ4K](https://perma.cc/GC8X-VQ4K)). In short, the Dat protocol is a peer-to-peer protocol (i.e., decentralized and openly accessible) that provides non-adjustable timestamps to each change that occurs within a folder, which is given a permanent unique address on the peer-to-peer Web (36^64 addresses possible) [@Dat2017]. The full details, implications, and potential implementations of this protocol for scholarly communication fall outside of the scope of this paper (see [https://dat-com-chris.hashbase.io](https://web.archive.org/web/20180316145316/https://dat-com-chris.hashbase.io/)).

@@

  With a chronological ordering of various modular research outputs and their parent relations, a directional adjacency matrix can be extracted for network analysis. Table 1 shows the directional adjacency matrix for Figure 1. Parent nodes (i.e., modules) must precede the child nodes in time, therefore only $\frac{J(J-1)}{2}$ of cells of the adjacency matrix are filled in, where $J$ is the number of research modules.

@@

  We propose to communicate research in modular "as-you-go" outputs (e.g., theory followed by hypotheses, etc.) instead of large "after-the-fact" papers. Modular communication opens up the possibility of a network of knowledge to come into existence when these pieces are linked (e.g., results descend from data). This network of knowledge would be supplementary to traditional citation networks and could facilitate new evaluation tools that are based in the question of interest than generic "one size fits all" evaluation indicators (e.g., Journal Impact Factor, citation counts, number of publications). Given the countless questions and operationalizations possible to evaluate research in a network of knowledge, we hope this would increase the focus on indicators as a tool in the evaluation process instead of indicators being the evaluation process itself [@Hicks2015].

@@

  Communicating scholarly research in modular "as-you-go" outputs might also address other threats to research sustainability. In modular "as-you-go" communication, selective publication based on results would be reduced because data would be communicated before results are known. Similarly, adjusting predictions after results are known would be reduced because predictions would be communicated before data are available (i.e., preregistration by design). Replications (or reanalyses) would be encouraged both for the replicated (the replicated node gets more child nodes, increasing its centrality) and the replicator (time investment is lower due to only having to add a data node that is linked to the materials node of the replicated). Self-plagiarism could be reduced by not forcing researchers to rehash the same theory across papers that spawn various predictions and studies. These various issues (amongst other out of scope issues) could be addressed jointly instead of each issue vying for importance for researchers, funders, or policy makers (amongst others).

  To encourage culture- and behavioral change, "after-the-fact" papers and modular "as-you-go" outputs could co-exist (initially) and would not require researchers to make a zero-sum decision. Copyright is often transferred to publishers upon publication (resulting in pay-to-access), but only after a legal contract is signed. Hence, preprints cannot be legally restricted by publishers when they precede a copyright transfer agreement. However, preprints face institutional and social opposition [@Kaiser2017], where preprinting could exclude a manuscript for publication depending on editorial policies or due to fears of non-publication or scooping (itself a result of hypercompetition). In recent years, preprints have become more widely accepted and less likely to exclude manuscript publication (e.g., Science accepts preprinted manuscripts) [@Berg2017]. Similarly, sharing modular "as-you-go" outputs would not be legally restricted by publishers and could ride the wave of preprint acceptance, but might also face institutional or social counterchange similar to the history of preprints. Researchers could communicate "as-they-go" and compile "after-the-fact" papers, facilitating co-existence and minimizing negative effects on career opportunities. Additionally, “as-you-go” modules could be used in scholarly fields where the provenance of information is important to findings.

  As far as we know, modular "as-you-go" scholarly communication infrastructure that includes direct links has not yet been available to researchers in a sustainable way. The only thought style that has facilitated "as-you-go" reporting in the past decade is that of Open Notebook Science (ONS) [@Bradley2007], where researchers share their day-to-day notes and thoughts. However, ONS has remained on the fringes of the Open Science thought style and has not matured, limiting its usefulness and uptake. For example, ONS increases user control because communication occurs on personal domains, but does not have a mechanism of preserving the content. Considering reference rot occurs in seven out of ten scholarly papers containing Weblinks [@Klein2014], concern for sustainable ONS is warranted without further development of content integrity. Moreover, ONS increases information output without providing more possibilities of discovering that content. Nanopublications have been implemented recently [@10.7717/peerj-cs.78].
